### 2024-08-18 paper.tex
remaining setup for finding Jordan basis in section 8

\section{Springer fibers}
Let \(G \subseteq \GL_m(\mathbb{C})\) be a connected semisimple Lie group, and let \(\mathfrak{g} \subseteq \gl_m(\mathbb{C})\) be its Lie algebra.
% ^I take this scope because it is about what Yuzhou Gu said.
Let \(\mathcal{N} \subseteq \mathfrak{g}\) be the subset consisting of nilpotent elements.
Let \(\mathcal{B}\) be the variety of Borel subalgebras of \(\mathfrak{g}\).
Let \(\widetilde{\mathcal{N}} = \{(\mathfrak{b}, n) : n \in \mathfrak{b}\} \subseteq \mathcal{B} \times \mathcal{N}\).
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N}\) be the projection onto the second coordinate.
For \(n \in \mathcal{N}\), we call \(\pi^{-1}(n)\) the \emph{Springer fiber at \(n\)}.
\par We mention some results about Springer fibers, which we will use later in this paper.
TODO: mention them.

@@

\section{Finding the irreducible components of \(U_0\)}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\), define 
\[U_{I, (\rho_i)} := \] 
\[\{(A_{0, b}, U) \in U_0 : \{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
\begin{lemma}
    \(U_0 = \bigcup_{I, (\rho_i)} U_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
    TODO: something about overlap being small (they're not disjoint).  
    They are distinct.
    Is that good enough for my purposes?
\end{lemma}
\begin{proof}
    Let \((A_{0, b}, U) \in U_0\).
    Let \(I' = \{(p_i, q_i) : p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I'\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

We note that even finding the irreducible components of \(\{A_{0,0}, U \in V\}\) is a pretty hard problem.
Not ambitious enough to try to re-solve that problem, we will use its solution.
\par To find the irreducible components of \(U_0\), we will use a known result about the irreducible components of the usual Springer fiber at an element \(J(\mu) \in \gl_{m + n}\).
(I should define this somewhere else, but \(J(\mu)\) just has the blocks in order of nonincreasing size.)
\begin{theorem}
    (Needs citation!)
    The irreducible components of the Springer fiber at \(J(\mu)\) are in bijection with the standard Young tableaus of shape \(\mu\).
    Further, the irreducible components are equidimensional, of dimension \(\sum_{i \neq j} \min(\mu_i, \mu_j) = \sum_i (i - 1) \mu_i\).
\end{theorem}

@@

We will use this result; to find the regular Springer fiber at any \(X \in \gl_{m + n}\), we just need to find \(P\) such that \(P^{-1}XP\) is in Jordan normal form; equivalently, we just need to find a Jordan basis for \(X\).
\par So, our strategy will be to find a subset of the possible values of \(b\) such that \(A_{0,b}\) has some fixed Jordan form \(\mu\), and we have an algebraic map \(b \mapsto P_b\) such that for \(A_{0,b}\) in the subset, \(P_bJ(\mu)P_b^{-1} = A_{0,b}\).
Then we obtain an irreducible component by picking one of these subsets of the possible values of \(b\) and picking a Springer fiber at \(J(\mu)\).
\par Finding the map \(b \mapsto P\) amounts to finding a Jordan basis for \(A_{0,b}\).
So, with this as our motivation, we find a Jordan basis for \(A_{0,b}\).

>>>>

\section{Springer fibers}
Let \(G \subseteq \GL_m(\mathbb{C})\) be a connected semisimple Lie group, and let \(\mathfrak{g} \subseteq \gl_m(\mathbb{C})\) be its Lie algebra.
% ^I take this scope because it is about what Yuzhou Gu said.
Let \(\mathcal{N} \subseteq \mathfrak{g}\) be the subset consisting of nilpotent elements.
Let \(\mathcal{B}\) be the variety of Borel subalgebras of \(\mathfrak{g}\).
Let \(\widetilde{\mathcal{N}} = \{(\mathfrak{b}, n) : n \in \mathfrak{b}\} \subseteq \mathcal{B} \times \mathcal{N}\).
Let \(\pi : \widetilde{\mathcal{N}} \to \mathcal{N}\) be the projection onto the second coordinate.
For \(n \in \mathcal{N}\), we call \(\pi^{-1}(n)\) the \emph{Springer fiber at \(n\)}.
\par We mention some results about Springer fibers, which we will use later in this paper.
\begin{theorem}
    (Needs citation!)
    The irreducible components of the Springer fiber at \(J(\mu)\) are in bijection with the standard Young tableaus of shape \(\mu\).
    Further, the irreducible components are equidimensional, of dimension \(\sum_{i \neq j} \min(\mu_i, \mu_j) = \sum_i (i - 1) \mu_i\).
\end{theorem}

@@

\section{Finding the irreducible components of \(U_0\)}
Let \(X\) be nilpotent with Jordan basis \((e_{ij}')_{1 \leq i \leq k, 1 \leq j \leq \lambda_i'}\).
Let \(V = \{(A_{a,b}, U) : \forall i. \; A_{a,b} U_{i + 1} \subseteq U\}\).
Let
\[U_0 = \{(A_{0,b}, U) \in V\}.\]
We will find the irreducible components of \(U_0\).
\par We write \(b_{ij}\) to denote the projection of \(b \in \mathbb{C}^m\) onto \(e_{ij}\).
For each row \(i\), let \(p_i(b) = \max\{j : b_{ij} \neq 0\}\) (the maximum of the empty set is zero).
Then set \(q_i(b) = \lambda_i - p_i(b)\).
When it is clear enough from context, we will just write \(p_i\) and \(q_i\) instead of \(p_i(b)\) and \(q_i(b)\).
\par For any \(I \subseteq \{1, ..., k\}\) and any values \((\rho_i)_{i \in I}\) such that (1) \(\rho_i \leq \lambda_i\), (2) \(\rho_i\) is decreasing with \(i\), and (3) \(\lambda_i - \rho_i\) is decreasing with \(i\), define 
\[U_{I, (\rho_i)} := \] 
\[\{(A_{0, b}, U) \in U_0 : \{(p_i, q_i) : i \in I\} = \{(p_i, q_i) : p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j \} \] 
\[\land \forall i \in I. \; p_i = \rho_i \}.\]
\begin{lemma}
    \(U_0 = \bigcup_{I, (\rho_i)} U_{I, (\rho_i)}\), where \(I\) ranges over all subsets of \(\{1, ..., k\}\), and \((\rho_i)\) ranges over all maps \(I \to \mathbb{N}_{>0}\) such that \(\rho_i\) and \(\lambda_i - \rho_i\) are both decreasing with \(i\).
    TODO: something about overlap being small (they're not disjoint).  
    They are distinct.
    Is that good enough for my purposes?
    Ah, I think it is good enough if I instead say that the sets of \(b\)'s are distinct.
\end{lemma}
\begin{proof}
    Let \((A_{0, b}, U) \in U_0\).
    Let \(I' = \{(p_i, q_i) : p_i = \max_{j : q_j = q_i} p_j > \max_{j : q_j < q_i} p_j\}\).
    Then let \(I = \{\max\{i : (p_i, q_i) = (p, q)\} : (p, q) \in I'\}\).
    Choosing the maximal \(i\) is an arbitrary choice; we just have to pick one.
    For \(i \in I\), \(\rho_i := p_i\).
    All we need to do is show that each \(\rho_i\) is positive, and that \(\rho_i\) and \(\lambda_i - \rho_i\) are decreasing.
    Then we get that \((A_{0,b}, U) \in U_{I, (\rho_i)}\), and we are done.
    \par Recall that by convention the maximum of the empty set is zero.
    So, the fact that \(p_i > \max_{j > i} p_j\) tells us that \(\rho_i = p_i\) is positive.
    \par Let \(i, j \in I\) with \(i < j\).
    Since \(\lambda_i \geq \lambda_j\), we have either \(p_i \geq p_j\) or \(q_i \geq q_j\).
    Further, since \((p_i, q_i) \neq (p_j, q_j)\) (clear from definition of \(I\)), either \(p_i > p_j\) or \(q_i > q_j\).
    And for \(i, j \in I'\) we clearly have \(p_i > p_j \iff q_i > q_j\).
    So in either case, both \(p_i > p_j\) and \(q_i > q_j\), and consequently both \(\rho_i > \rho_j\) and \(\lambda_i - \rho_i > \lambda_j - \rho_j\).
\end{proof}

\begin{lemma}
    Here we have an alternative characterization of \(U_{I, (\rho_i)}\).
    Namely, \(p_i = \rho_i\) for \(i \in I\) and for \(i \notin I\), we have \(p_i \leq something\).
\end{lemma}
\begin{proof}

\end{proof}
Fix any \(I\) and \((\rho_i)\) satisfying the conditions of Lemma 8.1.
We will find the irreducible components of \(U_{I, (\rho_i)}\), and show that their closures are in fact irreducible components of \(U_0\).
\par I claim that \(A_{0,b}\) has the same shape for every \((A_{0, b}, U) \in U_{I, (\rho_i)}\).
Call this shape \(\mu\).
By finding a algebraic map taking \(b\) to a Jordan basis for \(A_{0,b}\), we will put \(U_{I, (\rho_i)}\) in isomorphism with the product (choice of \(b\)) \(\times\) (Springer fiber at \(J(\mu)\)).
Then we will use our result about the usual Springer fiber at \(J(\mu)\) to find the irreducible components of \(U_{I, (\rho_i)}\).
(I should define this somewhere else, but \(J(\mu)\) just has the blocks in order of nonincreasing size.)

\par TODO: where does this paragraph go?

@@

\par TODO: begin Jordan basis thing in earnest here.
